Collaboratively Testing the Validity of Neuroscientific Models

### 2013-11-30 NeurounitFrontiers/Paper.tex
Added ipython reference and fixed one sentence

%In many cases, models require no modifications to take the new tests because the same type of model output is being requested.
\subsection{Implementing a Validation Test in \textit{SciUnit}}
Fig. \ref{fig:rate_test} shows how a scientist can implement spike count tests such as the one described above using \textit{SciUnit}. 
A \textit{SciUnit} validation test is an {instance} (i.e. an object) of a Python class implementing the \verbx{sciunit.Test} interface (cf. line 1). 
Here, we show a class \verbx{SpikeCountTest} taking three \emph{parameters} in its constructor (constructors are named \verbx{__init__} in Python, lines 9-10). 
The meaning of each parameter along with a description of the goodness-of-fit metric used by the test is documented on lines 4-7. 
%For convenience, we also make use of functions provided by the popular NumPy\cite{numpy_url} and SciPy\cite{scipy_url} libraries, although these are not required by \textit{SciUnit}.  
To create a \emph{particular} spike count test, we instantiate this class with particular experimental observations. 
For example, given observations from CA1 cells (not shown), we can instantiate a test as follows\todo{use iPython style In[0] notation?}:
\begin{ipy}
  In [0]: CA1_sc_test = SpikeCountTest(CA1_inputs, CA1_means, CA1_stds)
\end{ipy}
We emphasize the crucial distinction between the \textit{class} \verbx{SpikeCountTest}, which defines a \emph{parameterized family} of validation tests, and the particular \textit{instance} \verbx{CA1_sc_test}, which is an individual validation test because the necessary parameters, derived from data, have been provided. 
As we will describe below, we expect communities to build repositories of such families capturing the criteria used in their subfields of neuroscience so that test generation for a particular system of interest will often require simply instantiating a previously-developed family with particular experimental parameters and data. 
For single-neuron test families like \verbx{SpikeCountTest}, we have developed such a library, called \textit{NeuronUnit} (\url{http://github.com/scidash/neuronunit}) (Sec. \ref{sec:neuronunit}). Particular tests, like \verbx{CA1_sc_test}, will be derived from these families in suite repositories (e.g. \verbx{HippocampusTests}, as shown in Figure \ref{fig:sciunit_overview}).

@@

\subsection{Test Suites and Score Matrices}\label{sec:visualization} A collection of tests intended to be run on the same model can be put together to form a test suite.
The following test suite could be used for a simplified version of the QSNMC, as described in Sec. \ref{sec:example}:  
\begin{ipy}
  In [5]: CA1_suite = sciunit.TestSuite([CA1_overall_test, CA1_sc_test, CA1_sl_test, CA1_sv_test, CA1_isi_test, CA1_imi_test])
\end{ipy}
Like a single test, a test suite is capable of judging one or more models. The models must possess the union of the capabilities of the tests in a suite to be fully capable of running the suite, though partial results can also be generated. When a model cannot take a test because of a missing capability, the \texttt{NotApplicable} score is used (e.g. point processes cannot predict subthreshold voltages). 
The result is a \emph{score matrix}, like the one diagrammed in Fig. \ref{fig:sciunit_overview}.
\begin{ipy}
  In [6]: scores = CA1_suite.judge([ARX_model, AdEx1_model, aSRM_model, PP_model])
\end{ipy}
A simple summary of the scores in a score matrix can be printed to the console or visualized by other tools.  For example, the score matrix can be visualized as a hyperlinked table inside an IPython notebook \todo{reference to ipython nb docs/paper?} (Fig. \ref{fig:scidash_matrix}).

@@

NeuroConstruct (\cite{gleeson_neuroconstruct:_2007}, \url{http://www.neuroconstruct.org}) is a simulation manager that takes NeuroML models and hands off simulation to supported simulators. 
\textit{NeuronUnit} offers a \verbx{sciunit.Model} subclass called \verbx{NeuroConstructModel}, parameterized by a path to a NeuroML model.  
NeuroML can describe a wide range of models, but \verbx{NeuroConstructModel} currently supports only those that are time-integrable point neurons (exposing capabilities \verbx{TimeIntegrable} and \verbx{HasMembranePotential}). It can be instantiated directly or subclassed (useful if additional capabilities need to be manually exposed, for example) to generate a model suitable for testing (Fig. \ref{fig:ca1_model}).

@@

\subsection{A Complete Pipeline}
Although the tools described in Sec. \ref{sec:neuronunit} do not exhaust the possible sources of models, capabilities, and test data, they provide an immediate point of entry into the neurophysiology community and a powerful demonstration of our proposal. 
In the \textit{NeuronUnit} repository (\url{http://github.com/scidash/neuronunit}) is a runnable script (\textit{examples.py}) demonstrating a complete testing pipeline. 
It (1) selects an OSB model; 
(2) simulates it using NeuroConstruct; 
(3) tests the widths of the resulting action potentials, extracted and computed using NeuroTools, against NeuroElectro data downloaded on-the-fly, using a \textit{NeuronUnit} test class called \verbx{SpikeWidthTestDynamic}; \todo{Update this with current examples.py content}
and (4) computes and prints a test score (Figure~\ref{fig:neuronunit_example}).

>>>>

%In many cases, models require no modifications to take the new tests because the same type of model output is being requested.
\subsection{Implementing a Validation Test in \textit{SciUnit}}
Fig. \ref{fig:rate_test} shows how a scientist can implement spike count tests such as the one described above using \textit{SciUnit}. 
A \textit{SciUnit} validation test is an {instance} (i.e. an object) of a Python class implementing the \verbx{sciunit.Test} interface (cf. line 1). 
Here, we show a class \verbx{SpikeCountTest} taking three \emph{parameters} in its constructor (constructors are named \verbx{__init__} in Python, lines 9-10). 
The meaning of each parameter along with a description of the goodness-of-fit metric used by the test is documented on lines 4-7. 
%For convenience, we also make use of functions provided by the popular NumPy\cite{numpy_url} and SciPy\cite{scipy_url} libraries, although these are not required by \textit{SciUnit}.  
To create a \emph{particular} spike count test, we instantiate this class with particular experimental observations. 
For example, given observations from CA1 cells (not shown), we can instantiate a test as follows:
\begin{ipy}
  In [0]: CA1_sc_test = SpikeCountTest(CA1_inputs, CA1_means, CA1_stds)
\end{ipy}
We emphasize the crucial distinction between the \textit{class} \verbx{SpikeCountTest}, which defines a \emph{parameterized family} of validation tests, and the particular \textit{instance} \verbx{CA1_sc_test}, which is an individual validation test because the necessary parameters, derived from data, have been provided. 
As we will describe below, we expect communities to build repositories of such families capturing the criteria used in their subfields of neuroscience so that test generation for a particular system of interest will often require simply instantiating a previously-developed family with particular experimental parameters and data. 
For single-neuron test families like \verbx{SpikeCountTest}, we have developed such a library, called \textit{NeuronUnit} (\url{http://github.com/scidash/neuronunit}) (Sec. \ref{sec:neuronunit}). Particular tests, like \verbx{CA1_sc_test}, will be derived from these families in suite repositories (e.g. \verbx{HippocampusTests}, as shown in Figure \ref{fig:sciunit_overview}).

@@

\subsection{Test Suites and Score Matrices}\label{sec:visualization} A collection of tests intended to be run on the same model can be put together to form a test suite.
The following test suite could be used for a simplified version of the QSNMC, as described in Sec. \ref{sec:example}:  
\begin{ipy}
  In [5]: CA1_suite = sciunit.TestSuite([CA1_overall_test, CA1_sc_test, CA1_sl_test, CA1_sv_test, CA1_isi_test, CA1_imi_test])
\end{ipy}
Like a single test, a test suite is capable of judging one or more models. The models must possess the union of the capabilities of the tests in a suite to be fully capable of running the suite, though partial results can also be generated. When a model cannot take a test because of a missing capability, the \texttt{NotApplicable} score is used (e.g. point processes cannot predict subthreshold voltages). 
The result is a \emph{score matrix}, like the one diagrammed in Fig. \ref{fig:sciunit_overview}.
\begin{ipy}
  In [6]: scores = CA1_suite.judge([ARX_model, AdEx1_model, aSRM_model, PP_model])
\end{ipy}
A simple summary of the scores in a score matrix can be printed to the console or visualized by other tools.  For example, the score matrix can be visualized as a hyperlinked table inside an IPython notebook \citep{perez_ipython_2007} (Fig. \ref{fig:scidash_matrix}).

@@

NeuroConstruct (\cite{gleeson_neuroconstruct:_2007}, \url{http://www.neuroconstruct.org}) is a simulation manager that takes NeuroML models and hands off simulation to supported simulators. 
\textit{NeuronUnit} offers a \verbx{sciunit.Model} subclass called \verbx{NeuroConstructModel}, parameterized by a path to a NeuroML model.  
NeuroML can describe a wide range of models, but \verbx{NeuroConstructModel} currently supports only those that are time-integrable neurons with membrane potentials (exposing capabilities \verbx{TimeIntegrable} and \verbx{HasMembranePotential}). It can be instantiated directly or subclassed (useful if additional capabilities need to be manually exposed, for example) to generate a model suitable for testing (Fig. \ref{fig:ca1_model}).

@@

\subsection{A Complete Pipeline}
Although the tools described in Sec. \ref{sec:neuronunit} do not exhaust the possible sources of models, capabilities, and test data, they provide an immediate point of entry into the neurophysiology community and a powerful demonstration of our proposal. 
In the \textit{NeuronUnit} repository (\url{http://github.com/scidash/neuronunit}) is a runnable script (\textit{examples.py}) demonstrating a complete testing pipeline. 
It (1) selects an OSB model; 
(2) simulates it using NeuroConstruct; 
(3) tests the widths of the resulting action potentials, extracted and computed using NeuroTools, against NeuroElectro data downloaded on-the-fly, using a \textit{NeuronUnit} test class called \verbx{SpikeWidthTestDynamic}; \todo{Update this with current examples.py content, and changed to ipython notebook notation}
and (4) computes and prints a test score (Figure~\ref{fig:neuronunit_example}).
